Temporary Landing Page for Longitudinal Project

### 2023-04-20 11_Tutorials_MultivariateLatentGrowthCurveModels.qmd
tutorials update 1

---
title: "Multivariate Latent Growth Curves"
---

>>>>

---
title: "Multivariate Latent Growth Curves"
author: "Biostatistics Working Group"
---

## Overview
In this tutorial, we will introduce the Multivariate Latent Growth Curve Model (MLGCM), also known as Parallel Process Latent Growth Curve Model. These models allow us to analyze multiple growth processes simultaneously, examining the relationships between the initial levels and growth rates of multiple variables over time.

We will use the "lavaan" package in R to fit a MLGCM with two variables, each measured at three timepoints. We will also provide a brief overview of interpreting the results.

[+add diagrams/figures]

### When to use Multivariate Latent Growth Curve Models?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Multivariate Latent Growth Curve Models

## Basic Example
```{r}
# Load necessary packages
if (!("lavaan" %in% installed.packages())) install.packages("lavaan")
library(lavaan)

# Generate sample data
set.seed(42)
n <- 250
time1_var1 <- rnorm(n, mean = 10, sd = 3)
time2_var1 <- time1_var1 + rnorm(n, mean = 2, sd = 2)
time3_var1 <- time2_var1 + rnorm(n, mean = 2, sd = 2)

time1_var2 <- rnorm(n, mean = 5, sd = 2)
time2_var2 <- time1_var2 + rnorm(n, mean = 1, sd = 1)
time3_var2 <- time2_var2 + rnorm(n, mean = 1, sd = 1)

# Combine into a data frame
data <- data.frame(time1_var1, time2_var1, time3_var1,
                   time1_var2, time2_var2, time3_var2)
```

### Model Specification and Estimation
```{r}
# Specify the multivariate latent growth curve model
model <- '
  # Intercept and slope factors for variable 1
  i_var1 =~ 1*time1_var1 + 1*time2_var1 + 1*time3_var1
  s_var1 =~ 0*time1_var1 + 1*time2_var1 + 2*time3_var1

  # Intercept and slope factors for variable 2
  i_var2 =~ 1*time1_var2 + 1*time2_var2 + 1*time3_var2
  s_var2 =~ 0*time1_var2 + 1*time2_var2 + 2*time3_var2
'
```

```{r}
# Estimate the model
fit <- lavaan::sem(model, data = data)
```

```{r}
summary(fit, fit.measures = TRUE)
```

### Interpreting the Results
In the output, you will see the estimates for the factor loadings, intercepts, and slopes for both variables. These estimates describe the initial levels and growth rates for each variable. You will also see the variances and covariances of the latent intercept and slope factors, which provide information about individual differences in initial levels and growth rates, as well as the relationships between the initial levels and growth rates across the two variables.

Pay attention to the model fit indices (e.g., CFI, TLI, RMSEA, and SRMR) to evaluate how well the model fits the data. Good model fit is indicated by CFI and TLI values close to or greater than 0.95, RMSEA values close to or smaller than 0.06, and SRMR values close to or smaller than xxx

### 2023-04-20 15_Tutorials_RandomInterceptCrosslaggedPanelModels.qmd
tutorials update 1

---
title: "Random-Intercept Crosslagged Panel Models"
---

>>>>

---
title: "Random-Intercept Crosslagged Panel Models"
author: "Biostatistics Working Group"
---

## Overview
Random-Intercept Cross-Lagged Panel Models (RI-CLPM) are a type of structural equation model used to analyze longitudinal data with repeated measures of multiple variables. In this tutorial, we will walk you through a simple example using a simulated dataset and demonstrate how to fit a RI-CLPM using the `lavaan` package in R.

[+add diagrams/figures]

### When to use Random-Intercept Cross-Lagged Panel Models?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Random-Intercept Cross-Lagged Panel Models

#### Prerequisites
To follow along with this tutorial, you will need to have R and RStudio installed on your computer. Additionally, you will need to install the `lavaan` package, which provides functions for fitting various structural equation models, including RI-CLPMs.

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
# Install the lavaan package if not already installed
if (!("lavaan" %in% installed.packages())) {
  install.packages("lavaan")
}
# Load the lavaan package
library(lavaan)
```

## Basic Example 
Simulating the Data
For this tutorial, we will use a simulated dataset with 500 individuals, each measured at five time points. The dataset will contain two variables, X and Y, with cross-lagged effects.

```{r}
set.seed(123)
n <- 500
timepoints <- 5

# Simulate the data
data <- data.frame(id = rep(1:n, each = timepoints),
                   time = rep(1:timepoints, times = n),
                   X = rnorm(n * timepoints, mean = 0, sd = 1),
                   Y = rnorm(n * timepoints, mean = 0, sd = 1))

# Introduce cross-lagged effects
for (i in 2:timepoints) {
  data$X[data$time == i] <- 0.3 * data$Y[data$time == (i - 1)] + data$X[data$time == i]
  data$Y[data$time == i] <- 0.2 * data$X[data$time == (i - 1)] + data$Y[data$time == i]
}
```

### Model Specification and Estimation
Now that we have our simulated data, we can fit a Random-Intercept Cross-Lagged Panel Model using the lavaan package.

```{r, eval=FALSE}
# Define the model
model <- '
  # Random intercepts
  i_X =~ 1 * X_t1 + 1 * X_t2 + 1 * X_t3 + 1 * X_t4 + 1 * X_t5
  i_Y =~ 1 * Y_t1 + 1 * Y_t2 + 1 * Y_t3 + 1 * Y_t4 + 1 * Y_t5

  # Cross-lagged effects
  Y_t2 ~ beta_YX * X_t1
  Y_t3 ~ beta_YX * X_t2
  Y_t4 ~ beta_YX * X_t3
  Y_t5 ~ beta_YX * X_t4

  X_t2 ~ beta_XY * Y_t1
  X_t3 ~ beta_XY *
```

### Interpreting the Results
xxxxxx

### 2023-04-20 16_Tutorials_LatentCurveModelsStructuredResiduals.qmd
tutorials update 1

---
title: "Latent Curve Models with Structured Residuals"
---

>>>>

---
title: "Latent Curve Models with Structured Residuals"
author: "Biostatistics Working Group"
---

## Overview
In this tutorial, we will explore Latent Curve Models with Structured Residuals (LCM-SR). LCM-SR allows for the separation of between-person and within-person components of individual change over time. By using structured residuals, we can account for time-varying covariates and interactions between variables at both the within-person and between-person levels.

[+add diagrams/figures]

### When to use Latent Curve Models with Structured Residuals?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Latent Curve Models with Structured Residuals

#### Prerequisites
You need to have the `lavaan` and `semTools` packages installed in R to perform LCM-SR. Install them using the following commands:

```{r}
if (!("lavaan" %in% installed.packages())) install.packages("lavaan")
if (!("semTools" %in% installed.packages())) install.packages("semTools")
```

## Basic Example 
For this tutorial, we will use simulated data with 500 individuals and 5 timepoints. The data will include a time-varying covariate, covariate, and an outcome variable, outcome.

```{r}
set.seed(123)
library(lavaan)
library(semTools)

# Simulate data
n <- 500
timepoints <- 5

id <- rep(1:n, each = timepoints)
time <- rep(1:timepoints, n)

covariate <- rnorm(n * timepoints, mean = 0, sd = 1)
outcome <- rnorm(n * timepoints, mean = 0, sd = 1)

data <- data.frame(id, time, covariate, outcome)
```

### Model Specification and Estimation
To specify an LCM-SR, we will use the lavaan syntax to define the measurement model and the structural model. We will model the linear growth of the outcome variable while accounting for the effect of the time-varying covariate.

```{r}
model <- '
  # Latent variables
  intercept =~ 1 * outcome_t1 + 1 * outcome_t2 + 1 * outcome_t3 + 1 * outcome_t4 + 1 * outcome_t5
  slope =~ 0 * outcome_t1 + 1 * outcome_t2 + 2 * outcome_t3 + 3 * outcome_t4 + 4 * outcome_t5

  # Time-varying covariate effect on outcome
  outcome_t1 ~ c1 * covariate_t1
  outcome_t2 ~ c2 * covariate_t2
  outcome_t3 ~ c3 * covariate_t3
  outcome_t4 ~ c4 * covariate_t4
  outcome_t5 ~ c5 * covariate_t5

  # Residual variances and covariances
  outcome_t1 ~~ r1 * outcome_t1
  outcome_t2 ~~ r2 * outcome_t2
  outcome_t3 ~~ r3 * outcome_t3
  outcome_t4 ~~ r4 * outcome_t4
  outcome_t5 ~~ r5 * outcome_t5

  # Latent variable variances
  intercept ~~ i_var * intercept
  slope ~~ s_var * slope

  # Latent variable covariances
  intercept ~~ i_s_cov * slope
'
```

Model Estimation
Now we will estimate the model using the lavaan function sem().

```{r, eval=FALSE}
# Reshape data to wide format
wide_data <- spread(data, key = time, value = outcome, sep
```
### Interpreting the Results
xxxxx

### 2023-04-20 8_Tutorials_AutoregressiveCrosslaggedPanelModels.qmd
tutorials update 1

---
title: "Autoregressive Crosslagged Panel Models"
---

>>>>

---
title: "Autoregressive Crosslagged Panel Models"
author: "Biostatistics Working Group"
---

## Overview
xxxx
Autoregressive cross-lagged panel models (ACPMs) are a type of statistical model used to analyze longitudinal data and examine the temporal relationships between two or more variables over time. These models help to disentangle within-person effects from between-person effects and can estimate both autoregressive effects (influence of a variable on itself over time) and cross-lagged effects (influence of one variable on another variable over time).

autoregressive cross-lagged panel models, commonly known as CLPMs. CLPMs are a statistical method used to analyze longitudinal data, particularly when we want to examine the reciprocal relationships between two or more variables over time. In essence, CLPMs allow us to investigate whether one variable at an earlier time point can predict another variable at a later time point, and vice versa.

CLPMs are useful when we want to go beyond simple correlations between variables at different time points and instead model the dynamic interplay between them. Specifically, CLPMs account for the possibility that the relationship between two variables at one time point may influence their relationship at a subsequent time point. These models can also account for autocorrelation, meaning that the same variable measured over time is correlated with itself.

[+add diagrams/figures]

### When to use Autoregressive Cross-Lagged Panel Models (ACPMs)?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Autoregressive Cross-Lagged Panel Models (ACPMs)
In this tutorial, we will introduce the concept of autoregressive cross-lagged panel models (ACPMs) and guide you through a simple example using a larger dataset. By the end of this tutorial, you will be able to:

1. Understand the basic concepts of ACPMs.
2. Fit an ACPM using example data in R.
3. Interpret the results of the model.

## Basic Example
For this tutorial, we will use a simulated dataset containing the following information for a group of 100 individuals:

- Subject ID
- Time (in years)
- Variable 1 (e.g., stress level)
- Variable 2 (e.g., job satisfaction)

```{r}
set.seed(123)  # For reproducibility
n_subjects <- 100
n_timepoints <- 3
subject_ids <- factor(rep(1:n_subjects, each = n_timepoints))
time <- rep(0:(n_timepoints - 1), times = n_subjects)
variable1 <- rnorm(n_subjects * n_timepoints, mean = 3, sd = 1)
variable2 <- rnorm(n_subjects * n_timepoints, mean = 3, sd = 1)

data <- data.frame(subject_id = subject_ids, time = time, variable1 = variable1, variable2 = variable2)
```

### Model Specification and Estimation
We will use the lavaan package in R to fit an ACPM. First, install the required package, if not already installed:

```{r}
if (!("lavaan" %in% installed.packages())) install.packages("lavaan")
library(lavaan)
```
Define the ACPM with three timepoints:

```{r}
model <- 
  # Autoregressive paths
  variable1_t2 ~ a1 * variable1_t1
  variable1_t1 ~ a2 * variable1_t0
  variable2_t2 ~ a3 * variable2_t1
  variable2_t1 ~ a4 * variable2_t0

  # Cross-lagged paths
  variable1_t2 ~ b1 * variable2_t1
  variable1_t1 ~ b2 * variable2_t0
  variable2_t2 ~ b3 * variable1_t1
  variable2_t1 ~ b4 * variable1_t0

```
Fit the ACPM using the lavaan function:

```{r}
#fit <- lavaan(model, data = data, missing = )

```

### Interpreting the Results
xxxxxxxx

### 2023-04-20 index.qmd
rstudio test commit 3
tutorials update 1

---
title: "swhawes.github.io"
---

This is a Quarto website.

To learn more about Quarto websites visit <https://quarto.org/docs/websites>.

>>>>

---
title: "Temporary Landing Page for Longitudinal Project"
---

This is github pages test site for the longitudinal analysis project. The information included within the site is for testing purposes only and contents may be inaccurate.
